How to Avoid OOTA Without Really Trying

### 2024-09-11 paper/ootavacuous.tex
Further wordsmithing of abstract
Reported-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

\begin{abstract}
	Attempts to create memory models that forbid out-of-thin-air
	(OOTA) behaviors of C++ \co{memory_order_relaxed} accesses
	have been either non-executable, complex, or unloved by
	implementers.
	However, we know of no instances of OOTA behavior in real C++
	implementations.

	We focus on shared-memory programs and C++ implementations based
	on traditional compilers and hardware, including CPUs and GPGPUs,
	thus permitting easier analysis of semantic dependencies and
	OOTA behaviors, thanks to existing formal hardware models.
	We show that these models' constraints prevent OOTA cycles from
	occurring in undefined-behavior-free C++ programs running on
	compiler-based implementations, provided the cycles involve only
	volatile atomics.
	We also extend this work to nonvolatile atomics by defining
	``quasi-volatile'' behavior that we C++ implementations
	must adhere to in order to evaluate arithmetic expresssions
	involving atomic accesses in accordance with the abstract
	machine.

	It follows that such implementations need take no special action
	to avoid OOTA, and that no changes are required to user code or
	to the standard.
\end{abstract}
\maketitle{}

>>>>

\begin{abstract}
	Attempts to create memory models that forbid out-of-thin-air
	(OOTA) behaviors of C++ \co{memory_order_relaxed} accesses
	have been either non-executable, complex, or unloved by
	implementers.
	At the same time, we know of no instances of OOTA behavior in real C++
	implementations.

	We focus on shared-memory programs and C++ implementations based
	on traditional compilers and hardware, including CPUs and GPGPUs,
	thus easing analysis of semantic dependencies and OOTA behaviors
	via existing hardware models.
	We show that these models' constraints prevent OOTA cycles from
	occurring in undefined-behavior-free C++ programs running on
	compiler-based implementations that use single-thread analysis,
	provided the cycles involve only volatile atomics.
	We also extend this work to nonvolatile atomics by defining
	``quasi-volatile'' behavior that C++ implementations must adhere
	to in order to evaluate arithmetic expressions involving atomic
	accesses in accordance with the standard.

	It follows that enforcing OOTA avoidance requires no special
	action from implementations and no changes to user code or
	to the standard.
\end{abstract}
\maketitle{}
